Polynomial expansions of finite volume data in a cut cell context

### 2023-04-08 example/DomainDecomposition/EBCellMerge/hoeb_geometry_test/_doc/paper.tex
broad strokes of paper strategy outlined

\title{Polynomial expansions of finite volume data in a cut cell context}
\author{N. Overton-Katz \footnotemark[1]
   \and H. Johansen     \footnotemark[1]
   \and D. T. Graves    \footnotemark[1]
   \and D. Devendran    \footnotemark[1]
   \and O. Anteparra    \footnotemark[1]}

\maketitle
\footnotetext[1]{Lawrence Berkeley National Laboratory, Berkeley,
  CA. Research at LBNL was supported financially by the Office of
  Advanced Scientific Computing Research of the US Department of
  Energy under contract number DE-AC02-05CH11231.}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}\

@@

EB grids can produce cells that arbitrarily small.    This produces
algorithmic difficulties that generations of researchers have labored
to overcome with great success.   EB methods are used for high speed
compressible flows \cite{compresible_ns_paper, ebamrhscl} and for
slower, incompressible applications \cite{miller-trebotich}.

@@

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P} C^p (\xbar - \xbar)^p \label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac(1){p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}

In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms  moments ($m$)  which are the natural
products of grid generation.
Modern, higher order embedded boundary methods use this description
directly to generate stencils
 \cite{nate-stokes-paper,  dharshi-poisson, geometry-paper}.    At the core
of these algorithms is a similar procedure for finding the Taylor
coefficients $C$ which is
worth discussing.

\section{Computing Taylor coefficients}

Consider a finite volume grid where we know all geometric moments to
order $P$.   Given values of $<\psi>$ everwhere on the grid, we wish to
approximate the Taylor coefficients at a volume $\ibold$ given that we
know $\cn_\ibold = \{\jbold\}$, the set of $N_v$ closest neighbors of
the volume $\ibold$.   Using the Taylor expansion at each point gives
us $N_v$ equations for $C^p$
\begin{equation}
  V_\jbold <\psi>_\jbold = \sum\limits_{p < P} C^p_\ibold m^p_\jbold.
\end{equation}
Let's say there are $N_p$ Taylor coefficients.   We can force
$N_v > N_p$ by getting a large enough neighborhood.    This means that
we have an overdetermined system of the form
\begin{equation*}
M C = P 
\end{equation*}
where $M = \{\{m^{pi}_\jbold\}\}$ and $P= \{<\psi_\jbold>\}$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Diagonal weighing matricies are common.  Typically, the equations for volumes near the
target volume receive higher weights.   Since these matricies are
(deliberately) not square, we approximate the solution using a
Moore-Penrose pseudo-inverse:
\begin{equation*}
C = ((WM)^T(WM))^{-1} (WM)^T P.
\end{equation*}
The matrix inversion at the heart of this  method requires that we
need know to what degree the matrix $A = (WM)^T(WM)$  is invertable.

>>>>

\title{Polynomial expansions of finite volume data in a cut cell context}
\author{N. Overton-Katz \footnotemark[1]
   \and H. Johansen     \footnotemark[1]
   \and D. T. Graves    \footnotemark[1]
   \and D. Devendran    \footnotemark[2]
   \and O. Anteparra    \footnotemark[1]}

\maketitle

@@

EB grids can contain volumes that are arbitrarily small.    This
"small-cell problem'' produces
algorithmic difficulties that generations of researchers have labored
to overcome with great success.   EB methods are used for high speed
compressible flows \cite{compresible_ns_paper, ebamrhscl} and for
slower, incompressible applications \cite{miller-trebotich}.

@@

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P} C^p (\xbar - \xbar)^p \label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}

In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms moments ($m$) which are the natural
products of grid generation.  Modern, higher order embedded boundary
methods use this description directly to generate stencils
\cite{nate-stokes-paper, dharshi-poisson, geometry-paper}.  At the
core, all these algorithms have a similar procedure for finding the
Taylor coefficients $C$ which is worth discussing.

\section{Computing Taylor coefficients and Matrix Conditioning}

Consider a finite volume grid where we know all geometric moments to
order $P$.  Given values of $<\psi>$ everwhere on the grid, we wish to
approximate the Taylor coefficients at a volume $\ibold$ given that we
know $\cn_\ibold = \{\jbold\}$, the set of $N_v$ closest neighbors of
the volume $\ibold$.  Using the Taylor expansion at each point gives
us $N_v$ equations for $C^p$
\begin{equation}
  V_\jbold <\psi>_\jbold = \sum\limits_{p < P} C^p_\ibold m^p_\jbold.
\end{equation}
Let's say there are $N_p$ Taylor coefficients.  We can force $N_v >
N_p$ by getting a large enough neighborhood.  This means that we have
an overdetermined system of the form
\begin{equation*}
M C = P 
\end{equation*}
where $M = \{\{m^{pi}_\jbold\}\}$, $C=\{C^p\}$,  and $P= \{<\psi_\jbold>\}$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[3].  Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[4].
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since these matricies are (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation*}
C = ((WM)^T(WM))^{-1} (WM)^T P.
\end{equation*}
The matrix inversion at the heart of this method requires that we need
know to what degree the matrix $A \equiv (WM)^T(WM)$ is
well-conditioned.  Experience suggests that small cells play a role in
that conditioning and that the weighting matrix $W$ can be used to
mitigate conditioning issues that unweighted least squares may
produce.

To explore the roles of small cells and weighting functions, we first
introduce a merged cut cell space where there are no small cells.
This merged space still has geometric moments accurate to the same order 
as the original cut cell space.

We study how weighting matricies change the conditioning of $A$.  This is
done both with and without cell merger in the hope hope that this
exploration may lead to simpler, more robust cut cell algorithms to
solve partial differential equations at high order.

\section{Cell Merger}

Dammit I have to insert pictures here.

\footnotetext[1]{Lawrence Berkeley National Laboratory, Berkeley,
  CA. Research at LBNL was supported financially by the Office of
  Advanced Scientific Computing Research of the US Department of
  Energy under contract number DE-AC02-05CH11231.}
\footnotetext[2]{Ford Motor Company, Sunnyvale California.}
\footnotetext[3]{If the matrix $M$ is square, a weighting matrix can
  have no effect and the Moore-Penrose psueudoinverse becomes
  $M^{-1}$.}
\footnotetext[4]{Defining $\xbar_\ibold = 0$, the
  equations for volumes $\jbold$ where $\xbar_\jbold$ target receive
  higher weights.  Diagonal weighting matricies are common.  Usually,
  these algorithms define a distance metric $D(\ibold, \jbold)$ for
  two volumes $\ibold$ and $\jbold$.  Typically they make
  $W_{\jbold,\jbold}$ decrease strongly with increasing $D(\ibold,
  \jbold)$ to assign higher importance to the equations for volumes
  closer to $\ibold$.  Devendran, for example \cite{poisson-paper},
  uses a weighting function that varies with $W_{\ibold,\jbold}
  \approx 1./D(\ibold,\jbold)^5$ }.

\renewcommand{\thefootnote}{\fnsymbol{footnote}}\
